---
title: "Homework 4"
author: "Woo Young Choi"
date: "February 26, 2025"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Part 1: 

#### The template set up:
* fluidPage() - Creates the user interface
* sidebarLayout() - Organizes the UI into a sidebar and main panel
* sidebarPanel() - Where inputs will go
* mainPanel() - Where outputs will be displayed
* server function - Handles the app's logic
* shinyApp() - Combines the UI and server to create the application

#### Title Panel:

* Added a title to the app using titlePanel("k-means clustering"). This creates a header at the top of the page, giving users context about what the app does.

#### UI Inputs 
* input controls are added to the sidebarPanel():

#### UI Outputs
* Added plotOutput('plot1') to the mainPanel(). This creates a placeholder for where the plot will be displayed.

#### Output Plot 
* A simple plot is created in the server function using: 
```{r, eval=FALSE}
output$plot1 <- renderPlot({
  plot(iris[,c("Sepal.Length","Sepal.Width")])
})
```

#### Reactive Axes 
* Modified the plot to use the columns for the x and y axes 
```{r, eval=FALSE}
output$plot1 <- renderPlot({
  plot(iris[,c(input$xcol,input$ycol)])
})
```

#### K-means
* Added K-means clustering to the plot
```{r, eval=FALSE}
output$plot1 <- renderPlot({
  plot(iris[,c(input$xcol,input$ycol)],
       col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
})
```

#### Cluster Centers
* Enhanced the plot to show the cluster centers
```{r, eval=FALSE}
output$plot1 <- renderPlot({
  plot(iris[,c(input$xcol,input$ycol)],
       col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
  points(kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$centers, pch = 4, cex = 4, lwd = 4)
})
```

#### Save K-means 
* Improved the code by using a reactive expression 
```{r, eval=FALSE}
clusters <- reactive({
  kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)
})

output$plot1 <- renderPlot({
  plot(iris[,c(input$xcol,input$ycol)],
       col = clusters()$cluster)
  points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
})
```



### Part 2: Creating the K-means R Package

For this assignment, I made a comprehensive R package for K-means clustering called `myKMeansPackage`. The package has two main functions:

1. `kmeans_cluster()` - Performs K-means clustering on a dataset with proper input validation
2. `kmeans_plot()` - Visualizes the clustering results with customizable options

Both functions are shown with roxygen comments with examples. The package also includes comprehensive unit tests to correct error 


```{r, eval=FALSE}
library(myKMeansPackage)

# Load example data
data(iris)
iris_data <- iris[, 1:4]

# Perform clustering
result <- kmeans_cluster(iris_data, k = 3)

# Plot the results
kmeans_plot(iris_data, result, x_index = 1, y_index = 2)
```

The Shiny app is included in the package's `inst/shiny_app` directory and can be launched with:

```{r, eval=FALSE}
run_kmeans_app()
```

The benefits of the Shiny App

- Multiple data generation options (normal clusters, uniform random, circular pattern, or the iris dataset)
- Controls for the number of clusters and other K-means parameters
- Interactive visualization of clustering results
- Display of clustering statistics



### Part 3: GitHub Repository

My K-means package is available on GitHub at: [https://github.com/cwyjason/myKMeansPackage](https://github.com/cwyjason/myKMeansPackage)

The repository includes:
- Full source code for the package
- Documentation generated with roxygen2
- Unit tests created with testthat
- A comprehensive README file
- The Shiny app in the inst/shiny_app directory
